R-LIME: Rectangular Constraints and Optimization for Local Interpretable Model-agnostic Explanation Methods

### 2024-02-14 paper.tex
Complete 1st version

% settings for cleveref
\crefname{figure}{Fig.}{Figs.}
\crefname{table}{Tab.}{Tabs.}

@@

\section{Introduction}
Machine learning models, such as deep learning and random forests,
have been widely employed in various industrial applications
due to their significant improvement in accuracy in recent years.
However,
the increasing complexity and black-box nature of these models pose challenges,
particularly in critical decision-making scenarios like healthcare and finance,
where the opacity of decision rationale becomes a major obstacle to
implementation.
Consequently,
there has been extensive research in the field of post hoc explanation
for machine learning models \cite{%
  ribeiro2016why,ribeiro2018anchors,radulovic2023bella,guidotti2018local}.

Existing methods on post hoc explanations categorize into model-dependent and
model-agnostic methods,
based on their dependence on the model's structure.
Model-dependent methods primarily focus on deep neural networks and
explain the models' behavior using their parameters.
While they allow a direct approach to the model,
applying the same method to models with different structures is often challenging.
On the other hand,
model-agnostic methods do not rely on the models' parameters and utilize only their output.
Although they are applicable to any model, available information is limited.
Model-agnostic methods further classify into global and local methods
based on their treatment of input space.
Global methods aim to explain the model's behavior across the entire input space,
seeking explanations valid for any input.
However, providing global explanations becomes challenging as the model's complexity increases.
Local methods, in contrast, explain the model's output for a specific input
(focus point) in the vicinity of that input.
While local methods offer more simple and accurate explanations than global methods,
the scope of the explanation is limited to the vicinity of the focus point.

This paper specifically focuses on model-agnostic and local methods. For instance, Local Interpretable Model-agnostic Explanations (LIME) \cite{ribeiro2016why} approximates the complex model's decision boundary locally using a simpler model. Although LIME enables users to interpret the local behavior of complex models, it has been criticized for not optimizing the approximation region \cite{radulovic2023bella} and not presenting the approximation region in an interpretable form \cite{ribeiro2018anchors}.

To address the limitations of LIME,
we propose an algorithm that interprets complex models
in an interpretable approximation region,
taking inspiration from another explanation method called Anchor
\cite{ribeiro2018anchors}.
Our proposed method maximizes the accuracy of the learned approximation model
within a rectangular approximation region under lower-bound constraints.
The resulting rectangular region is expressed as a conjunction of feature predicates,
enhancing interpretability for the user.

\section{Related Work}
The figures presented visually compare three methods: LIME, Anchor, and BELLA.

\begin{figure}[t]
  \vspace{0.5cm}
  \begin{subfigure}[t]{\textwidth}
    \centering
    \includegraphics[width=0.55\textwidth]{lime_vs_anchor_exp_a.png}
    \caption{Instance around the focal point.}\label{fig:lime_vs_anchor_exp_a}
    \vspace{0.5cm}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{\textwidth}
    \centering
    \includegraphics[width=0.4\textwidth]{lime_vs_anchor_exp_b.png}
    \caption{Explanation by LIME.}\label{fig:lime_vs_anchor_exp_b}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{\textwidth}
    \centering
    \includegraphics[width=0.55\textwidth]{lime_vs_anchor_exp_c.png}
    \caption{Explanation by Anchor.}\label{fig:lime_vs_anchor_exp_c}
  \end{subfigure}
  \caption[Examples of output results by LIME and Anchor]{Examples of output results by LIME and Anchor \cite{ribeiro2018anchors}.}\label{fig:lime_vs_anchor_exp}
\end{figure}
In this chapter, we provide an overview of existing research on model-agnostic and local post hoc explanations, particularly focusing on studies closely related to the proposed method.

\subsection{LIME (Local Interpretable Model-agnostic Explanations) \cite{ribeiro2016why}}
LIME is a method for locally approximating a trained black-box classifier $f: \mathbb{R}^d \to \{0,1\}$ around a focal point $x \in \mathbb{R}^d$ using a linear model $g: \mathbb{R}^d \to \{0,1\}$ (\cref{fig:lime}). The process involves:
\begin{enumerate}
  \item Obtaining a set of perturbed samples $\mathcal{Z}_p$ around $x$ and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
  \item Learning an interpretable model $g$ using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$ by minimizing the following loss function:
        \begin{equation}
          \label{eq:lime_loss}
          \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
          \pi_x(z){\left(f(z)-g(z)\right)}^2,
        \end{equation}
        where $\pi_x(z)$ is a weight function designed to be larger for samples closer to $x$, typically implemented using an exponential kernel.
\end{enumerate}

LIME provides valuable insights into the local behavior of the model by showing the contribution of each feature to the output $f(x)$. However, due to not explicitly indicating the perturbation region, users cannot assess the effective scope of the explanation \cite{ribeiro2018anchors}. An example explanation for an LSTM sentiment prediction model using LIME is illustrated in \cref{fig:lime_vs_anchor_exp_b}. The left explanation suggests that the word "not" contributes to the model's positive prediction, which does not apply to the instance on the right. However, without the explicit perturbation region, users might mistakenly apply the left explanation to the right instance, potentially leading to a misunderstanding of the black-box model's behavior \cite{ribeiro2018anchors}.

\subsection{Anchor \cite{ribeiro2018anchors}}\label{sec:anchor}
Anchor represents a rectangular region,
expressed as a conjunction of feature predicates (rules),
that maximizes the probability of the black-box classifier $f$ outputting $f(x)$ within the region. It aims to highlight important features contributing significantly to the output (\cref{fig:anchor}).

@@

where $\mathcal{D}(z|A)$ is the conditional distribution in the region where the rule $A$ is true.
% When sampling the perturbation vector $z$ according to the distribution $\mathcal{D}$,
$\Prec(A)$ represents the probability that the output of $f$ matches between the perturbation $z$ and the focal point $x$ in the region $A$, and $\Cov(A)$ expresses the probability that the perturbation $z$ fits into $A$.

An Anchor maximizes coverage within the given threshold $\tau$ while maintaining accuracy.
However, direct computation of \cref{eq:anchor_def_prec} is not feasible, and it is evaluated using PAC (Probably Approximately Correct) through sampling.
Introducing a confidence level $1-\delta$ $(0\le\delta\le1)$, the accuracy constraint is relaxed as follows:
\begin{equation}
  \label{eq:const_prec}
  P(\Prec(A)\ge\tau)\ge1-\delta.
\end{equation}
Thus, the following optimization problem is solved:
\begin{equation}
  \label{eq:main_problem}
  A^*=\underset{A \text{~s.t. } P(\Prec(A)\ge\tau)\ge1-\delta\wedge A(x)=1}
  {\arg\max}\Cov(A).
\end{equation}

While Anchor provides a clear presentation of the effective scope of the explanation, the information it includes is less than LIME, limiting its utility \cite{ribeiro2018anchors}. An example explanation for an LSTM sentiment prediction model using Anchor is depicted in \cref{fig:lime_vs_anchor_exp_c}. The left explanation suggests that changing words other than "not" and "bad" has little impact on the classifier's output. While it clearly does not apply to the right instance, the explanation does not provide details about the influence of words like "not" or "bad," resulting in less user insight into the model's behavior compared to LIME (\cref{fig:lime_vs_anchor_exp_b}).

\subsection{BELLA (Black box model Explanations by Local Linear Approximations) \cite{radulovic2023bella}}
BELLA addresses the non-optimization of the approximation region, a drawback of LIME, by learning a local linear approximation model $g$ using a subset $Z'$ of the dataset $Z$ instead of perturbed samples (\cref{fig:bella}). The process involves:
\begin{enumerate}
  \item Computing the distance $d_x(z)$ for all instances $z$ in $Z$ with respect to the focal point $x$.
  \item Selecting the $k$ instances with the smallest $d_x(z)$, forming a subset $Z_s(k) \subseteq Z$.
  \item Learning the local linear approximation model $g$ using $Z_s(k)$ and calculating the similarity measure $\mathcal{R}(k)$ between $f$ and $g$.
  \item Obtaining the optimal $k^*$ by maximizing $\mathcal{R}(k)$.
\end{enumerate}

BELLA aims to enhance the reliability of explanations by optimizing the approximation region for high-accuracy linear approximations. However, it assumes direct access to the dataset $Z$ used for training the classifier, making it inapplicable in scenarios where direct access to the dataset is restricted for privacy reasons. Additionally, as BELLA presents the approximation region as a subset of the dataset, its interpretability is lower compared to Anchor.

\section{Proposed Method}
\begin{figure}[tb]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime}
    \caption{%
      LIME: Locally approximates the decision boundary around the focal point.}\label{fig:lime}
  \end{subfigure}%
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{anchor}
    \caption{Anchor: Maximizes coverage of a rectangular region containing the focal point under accuracy constraints.}\label{fig:anchor}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=0.85\textwidth]{bella}
    \caption{%
      BELLA: Explores a subset of the dataset, maximizing similarity between the local linear approximation model and the black-box classifier.}\label{fig:bella}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{rlime3}
    \caption{%
      R-LIME: Maximizes coverage of a rectangular region containing the focal point under lower constraints on the accuracy of the approximation model.}\label{fig:rlime}
  \end{subfigure}
  \caption{%
    Visual comparison of LIME, Anchor, BELLA and R-LIME (our proposed method).
    The solid line represents the rectangular region containing the focal
    point, and the dashed line represents the learned approximation model.
  }
\end{figure}

\subsection{Overview}
We propose R-LIME,
a method that aims to address the limitations of LIME \cite{ribeiro2016why}
and its improvement methods such as
Anchor \cite{ribeiro2018anchors} and BELLA \cite{radulovic2023bella}.
Similar to LIME,
the proposed method locally approximates the given black-box classifier $f$
around the focal point $x$ by a linear model $g$.
However,
by adopting a rectangular perturbation region similar to Anchor,
it represents feature predicates (rules).
The accuracy of explanations is defined as the "accuracy" of rules,
and the "coverage" of rules defines the generality of explanations.
The proposed method maximizes coverage under lower constraints on accuracy
(\cref{fig:rlime}).

Anchor maximizes the coverage of region $A$ within which the output of the black-box classifier $f$ matches $f(x)$ with respect to the probability under lower constraints. The proposed method, on the other hand, learns an approximation model $g$ within the rectangular region $A$ and maximizes the coverage of $A$ under lower constraints on the accuracy of $g$. We modify the definition of accuracy using \cref{eq:anchor_def_prec} as follows:
\begin{equation}
  \Prec(A)=\underset{g\in G}{\max}\ \mathbb{E}_{z\sim\mathcal{D}(z|A)}
  [\mathbbm{1}_{f(z)=g(z)}]. \label{eq:def_prec}
\end{equation}
where $G$ represents the set of possible linear approximation models. By solving the optimization problem in \cref{eq:main_problem} under this modified accuracy definition, we can select rules that learn high-accuracy approximation models, among which the coverage is maximized. The proposed method's algorithm, shown in \cref{sec:alg}, is largely based on the algorithms used in Anchor.

In this paper, we refer to the proposed method as "R-LIME" (Ruled LIME). The characteristics of R-LIME are listed below:
\begin{itemize}
  \item The perturbation vector's generation region is optimized and presented in a user-interpretable format.
  \item Interpretation is conducted without direct access to the dataset, utilizing only the distribution.
  \item The number of samples is dynamically determined based on the estimated accuracy without the need for a predefined amount.
\end{itemize}
% This enables model interpretation without direct access to the data, even in scenarios with significant privacy constraints. The optimized and presented explanation's applicable range allows users to evaluate the reliability and generality of the generated explanations.

\subsection{Algorithm}\label{sec:alg}
{%
  \begin{figure}[t]
    \centering
    \includegraphics[width=0.32\textwidth]{rlime1}
    \includegraphics[width=0.32\textwidth]{rlime2}
    \includegraphics[width=0.32\textwidth]{rlime3}
    \caption{%
      Overview of the R-LIME algorithm.
      The progression of the algorithm from left to right is illustrated.
      The solid line represents the rectangular region $A$,
      and the dashed line represents the linear approximation model $g$
      learned within $A$.
      The initial value of $A$ is an empty rule (entire input space),
      and predicates are added to $A$, reducing coverage.
      The process continues until $\Prec(A)\ge\tau$,
      at which point the rule with the maximum coverage is output.
    }
  \end{figure}
  \begin{algorithm}[p]
    \input{alg1}
  \end{algorithm}
  \begin{algorithm}[p]
    \input{alg2}
  \end{algorithm}
  \def\myidt{\hspace{\algorithmicindent}}
  \begin{algorithm}[p]
    \input{alg3}
  \end{algorithm}
  \begin{algorithm}[p]
    \input{alg4}
  \end{algorithm}
}
For non-convex optimization problems like \cref{eq:main_problem}, greedy methods are often employed. However, greedy methods tend to converge to local optima, and to address this, the proposed method utilizes beam search, which selects multiple candidates at each iteration. The pseudocode is shown in \cref{alg:greedy_search}.

\subsubsection{Candidate Rule Generation}
To generate new candidate rules, one additional predicate is added to each of the $B$ candidate rules obtained in the previous iteration. The pseudocode is shown in \cref{alg:generate_cands}. $T(x)$ represents the set of attribute-value pairs (predicates) that $x$ possesses, and $T(x)\setminus A$ represents the conjunction of predicates in $T(x)$ not included in rule $A$.

\subsubsection{Selection of Candidate Rule with Maximum Accuracy}
Given the set of generated candidate rules $\cands$, the algorithm selects the top $B$ candidates with the highest accuracy. This is treated as an optimal arm identification problem in the multi-armed bandit framework. Each candidate rule $A_i\in\cands$ is considered an arm, and their accuracy $\Prec(A_i)$ is treated as the distribution of rewards. By sampling $z\sim\mathcal{D}(\cdot|A_i)$ and obtaining the reward $\mathbbm{1}_{f(z)=g_i(z)}$ for each trial, the algorithm updates $g_i$ using the sampled perturbation vector $z$ and the pseudo-label $f(z)$ immediately after each trial. To efficiently select the rule (arm) with the highest accuracy, the KL-LUCB algorithm \cite{kaufmann2013information} is employed. The pseudocode is shown in \cref{alg:best_cands}. The KL-LUCB algorithm assumes that the reward distribution for each arm remains unchanged, while the proposed method updates the approximation model $g_i$ with each sampling, which may not satisfy this assumption. This issue is discussed further in \cref{sec:reward}.

\subsubsection{Selection of the Rule with Maximum Coverage Meeting the Constraint}
To satisfy the constraint imposed by \cref{eq:const_prec}, rule $A$ needs to fulfill the lower bound constraint:
\begin{equation}
  \Prec_{l}(A,\delta)>\tau
  \label{eq:stop_condition}
\end{equation}
where $\Prec_{l}(A,\delta)$ represents the lower limit of the $100(1-\delta)$\% confidence interval for $\Prec(A)$.
If the received set of candidate rules, $\mathcal{A}$, includes a rule satisfying \cref{eq:stop_condition},
the one with the maximum coverage among them is selected, and the iteration is concluded.
If $\mathcal{A}$ does not contain any rule satisfying \cref{eq:stop_condition}, it returns \textbf{null},
and proceeds to the next iteration.
The pseudocode is presented in \cref{alg:largest_valid_cand}.

The algorithmic procedure outlined above approximates the solution to the optimization problem in \cref{eq:main_problem} under the definition of accuracy in \cref{eq:def_prec}. It iteratively selects a rectangular region (rule) with the learned approximation model's accuracy exceeding a given threshold while maximizing coverage. The selected rule, with the maximum coverage meeting accuracy constraints, is output after each iteration.

\section{Experiments}
To verify the effectiveness of the proposed method, LIME and R-LIME were compared using a single dataset.

@@

The experiments utilized the recidivism dataset \cite{schmidt1988predicting}. This dataset contains information on 9,549 prisoners released from North Carolina prisons between July 1, 1979, and June 30, 1980. The dataset includes 19 items such as race, gender, presence of alcohol dependence, number of prior offenses, and recidivism. For this experiment, we treated the binary classification problem of predicting the presence or absence of recidivism (Recidivism) as the target label. Preprocessing included discretization of continuous features, resulting in 15 features. This problem setting simulates a case where a machine learning model is introduced to decide parole for prisoners. Since such decisions can have a significant impact on a person's life, it is crucial for users to interpret the outputs of black-box models appropriately.

The dataset, after removing missing values, was split into training data (7,639 instances) and test data (955 instances). A random forest model with 50 trees was trained using the training data. Subsequently, LIME and R-LIME explanations were generated for two instances extracted from the test data (\cref{fig:instance}). In R-LIME, logistic regression was used as the linear approximation model, and the distribution $\mathcal{D}$ was estimated using a multivariate normal distribution based on the training data.

The beam width was set to $B=10$, the confidence coefficient to $1-\delta=0.95$, and the tolerance of the KL-LUCB algorithm to $\epsilon=0.05$. Accuracy thresholds $\tau$ were set to $\tau=0.70,0.80,0.90$.

>>>>

% settings for cleveref
\crefname{figure}{Fig.}{Figs.}
\crefname{table}{Table.}{Tables.}

@@

\section{Introduction}
Machine learning models, such as deep learning and random forests,
have been widely employed in various industrial applications
due to their significant improvement in accuracy in recent years.
However,
the increasing complexity and black-box nature of these models pose challenges,
particularly in critical decision-making scenarios like healthcare and finance,
where the opacity of decision rationale becomes a major obstacle to
implementation.
Consequently,
there has been extensive research in the field of post hoc explanations
for machine learning models \cite{%
  ribeiro2016why,ribeiro2018anchors,radulovic2023bella,guidotti2018local}.

Existing methods on post hoc explanations categorize into
\emph{model-dependent} and \emph{model-agnostic} methods,
based on their dependence on the model's structure.
Model-dependent methods primarily focus on deep neural networks and
explain the models' behavior using their parameters.
While they allow a direct approach to the model,
applying the same method to models with different structures is often challenging.
On the other hand,
model-agnostic methods do not rely on the models' parameters and utilize only their output.
Although they are applicable to any model, available information is limited.
Model-agnostic methods further classify into
\emph{global} and \emph{local} methods based on their locality in input space.
Global methods aim to explain the model's behavior across the entire input space,
seeking explanations valid for any input.
However, providing global explanations becomes challenging as the model's complexity increases.
Local methods, in contrast, explain the model's output for a specific input
(focal point) in the vicinity of that input.
While local methods offer more simple and accurate explanations than global methods,
the scope of the explanation is limited to the vicinity of the focus point.

This paper specifically focuses on \emph{local} \emph{model-agnostic} methods.
For instance, LIME (Local Interpretable Model-agnostic Explanations)
\cite{ribeiro2016why} approximates the complex model's decision boundary
locally using a simpler model.
Although LIME enables users to interpret the local behavior of complex models,
it does not optimize the approximation region \cite{radulovic2023bella} and
the effective scope of the explanation is not explicitly presented
\cite{ribeiro2018anchors}.

To address the limitations of LIME,
we propose a new method called R-LIME (Ruled LIME),
which interprets complex classifiers in an interpretable region.
R-LIME locally approximates a complex decision boundary linearly
in a rectangular region and maximizes the region as long as the accuracy of the linear
classifier is higher than a specified threshold.
The resulting rectangular region is interpretable for users because it is
expressed as a conjunction of feature predicates.

\section{Related Work}
\begin{figure}[t]
  \centering
  \vspace{0.5cm}
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_a.png}
    \caption{Instances around the focal point.}\label{fig:lime_vs_anchor_exp_a}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_b.png}
    \caption{%
      Explanations by LIME.
      They provide contributions of each feature to the output,
      but do not explicitly indicate its scope.
    }\label{fig:lime_vs_anchor_exp_b}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_c.png}
    \caption{%
      Explanations by Anchor.
      They provide the effective scope of the explanation,
      but do not provide details about the influence of each feature.
    }\label{fig:lime_vs_anchor_exp_c}
  \end{subfigure}
  \caption{%
    Explanations by LIME and Anchor
    for an LSTM sentiment prediction model \cite{ribeiro2018anchors}.
  }\label{fig:lime_vs_anchor_exp}
\end{figure}
In this chapter,
we provide an overview of existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.

\subsection{LIME (Local Interpretable Model-agnostic Explanations) \cite{ribeiro2016why}}
LIME locally
approximates a trained black-box classifier $f: \mathbb{R}^d \to \{0,1\}$
around a focal point $x \in \mathbb{R}^d$
by a linear classifier $g: \mathbb{R}^d \to \{0,1\}$
(\cref{fig:lime}).
The process involves:
\begin{enumerate}
  \item Obtaining a set of perturbed samples $\mathcal{Z}_p$ around $x$
        and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
  \item Learning a linear classifier $g$
        using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
        by minimizing the following loss function:
        \begin{equation}
          \label{eq:lime_loss}
          \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
          \pi_x(z){\left(f(z)-g(z)\right)}^2,
        \end{equation}
        where $\pi_x(z)$ is a weight function designed to be larger for samples
        closer to $x$, typically implemented using an exponential kernel.
\end{enumerate}

LIME provides valuable insights into the local behavior of the model
by showing the contribution of each feature to the output $f(x)$.
However, due to not explicitly indicating the perturbation region,
users cannot assess the effective scope of the explanation
\cite{ribeiro2018anchors}.
An example of explanation by LIME for an LSTM sentiment prediction model
is illustrated in \cref{fig:lime_vs_anchor_exp_b}.
The left explanation suggests that
the word "not" contributes to the model's positive prediction,
which does not apply to the instance on the right.
However, without the explicit perturbation region,
users might mistakenly apply the left explanation to the right instance,
potentially leading to a misunderstanding of the black-box model's behavior
\cite{ribeiro2018anchors}.

\subsection{Anchor \cite{ribeiro2018anchors}}\label{sec:anchor}
Anchor represents a rectangular region containing the focal point $x$,
expressed as a conjunction of feature predicates (a rule),
that maximizes the probability of the black-box classifier $f$
outputting $f(x)$ within the region.
It aims to highlight important features
contributing significantly to the output (\cref{fig:anchor}).

@@

where $\mathcal{D}(z|A)$ is the conditional distribution in the region
where the rule $A$ returns true.
$\Prec(A)$ represents the probability that the output of $f$ matches
between the perturbation $z\sim\mathcal{D}(z|A)$ and the focal point $x$,
and $\Cov(A)$ expresses the probability that the perturbation $z$ fits into $A$.

Anchor maximizes coverage under the constraint that
the accuracy of the rule $A$ exceeds a given threshold $\tau$.
However, \cref{eq:anchor_def_prec} is not directly computable.
Introducing a confidence level $1-\delta$ $(0\le\delta\le1)$,
the accuracy constraint is relaxed as follows:
\begin{equation}
  \label{eq:const_prec}
  P(\Prec(A)\ge\tau)\ge1-\delta.
\end{equation}
Thus, the following optimization problem is solved:
\begin{equation}
  \label{eq:main_problem}
  A^*=\underset{A \text{~s.t. } P(\Prec(A)\ge\tau)\ge1-\delta\wedge A(x)=1}
  {\arg\max}\Cov(A).
\end{equation}

While Anchor provides an effective scope of the explanation,
the information it includes is less than LIME,
limiting its utility \cite{ribeiro2018anchors}.
An example of Anchor explanation for an LSTM sentiment prediction model
is depicted in \cref{fig:lime_vs_anchor_exp_c}.
The left explanation suggests that changing words other than "not" and "bad"
has little impact on the classifier's output.
While it clearly does not apply to the right instance,
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME
(\cref{fig:lime_vs_anchor_exp_b}).

\subsection{%
  BELLA (Black box model Explanations by Local Linear Approximations)
  \cite{radulovic2023bella}
}
BELLA addresses the non-optimization of the approximation region,
a drawback of LIME,
by learning a local linear classifier $g$
using a subset $Z'$ of the dataset $Z$
instead of perturbed samples (\cref{fig:bella}).
The process involves:
\begin{enumerate}
  \item Computing the distance $d_x(z)$ for all instances $z$ in $Z$
        with respect to the focal point $x$.
  \item Selecting the $k$ instances with the smallest $d_x(z)$,
        forming a subset $Z_s(k) \subseteq Z$.
  \item Learning the local linear classifier $g$ using $Z_s(k)$
        and calculating the similarity measure $\mathcal{R}(k)$ between $f$ and $g$
        over $k=1,2,\dots,|Z|$.
  \item Obtaining the optimal $k^*$ by maximizing $\mathcal{R}(k)$.
\end{enumerate}

BELLA aims to enhance the reliability of explanations
by optimizing the approximation region for high-accuracy linear approximations.
However, it assumes direct access to the dataset $Z$ used for training the classifier,
making it inapplicable in scenarios where direct access to the dataset is restricted.
Additionally,
as BELLA presents the approximation region as a subset of the dataset,
its interpretability is lower compared to Anchor.

\section{Proposed Method}
\begin{figure}[tb]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime}
    \caption{%
      LIME: Locally approximates the decision boundary around the focal point.
    }\label{fig:lime}
  \end{subfigure}%
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{anchor}
    \caption{%
      Anchor: Maximizes coverage of a rectangular region
      containing the focal point under accuracy constraints.
    }\label{fig:anchor}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=0.85\textwidth]{bella}
    \caption{%
      BELLA: Explores a subset of the dataset,
      maximizing similarity between the local linear classifier
      and the black-box classifier.
    }\label{fig:bella}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{rlime3}
    \caption{%
      R-LIME:
      Maximizes coverage of a rectangular region containing the focal point
      under lower constraints on the accuracy of the linear classifier.
    }\label{fig:rlime}
  \end{subfigure}
  \caption{%
    Visual comparison of LIME, Anchor, BELLA and R-LIME (our proposed method).
    The solid line represents the rectangular region containing the focal
    point, and the dashed line represents the learned approximation model.
  }
\end{figure}

\subsection{Overview}
We propose R-LIME,
a method that aims to address the limitations of LIME \cite{ribeiro2016why},
Anchor \cite{ribeiro2018anchors} and BELLA \cite{radulovic2023bella}.
Our method locally approximates the given black-box classifier $f$
around the focal point $x$ by a linear classifier $g$ similar to LIME and BELLA,
but samples parturbation from a rectangular region similar to Anchor
(\cref{fig:rlime}).
The accuracy of $g$ is defined as the "accuracy" of the rule,
and the generality of explanations is defined as the "coverage" of the rule.

Anchor maximizes the coverage of region $A$
as long as the probability of the output of the black-box classifier $f$ matching $f(x)$
within $A$ exceeds a given threshold $\tau$.
The proposed method, on the other hand, learns an linear classifier $g$
within the rectangular region $A$ and maximizes the coverage of $A$
under lower constraints on the accuracy of $g$. 
We modify the Anchor's definition of accuracy in \cref{eq:anchor_def_prec}
as follows:
\begin{equation}
  \Prec(A)=\underset{g\in G}{\max}\ \mathbb{E}_{z\sim\mathcal{D}(z|A)}
  [\mathbbm{1}_{f(z)=g(z)}]. \label{eq:def_prec}
\end{equation}
where $G$ is the set of possible linear classifiers.
By solving the optimization problem in \cref{eq:main_problem}
under the modified accuracy definition in \cref{eq:def_prec},
we can select the rule that enables explanation with high accuracy and generality.
The characteristics of R-LIME are listed below:
\begin{itemize}
  \item The approximation region is optimal and interpretable.
  \item The dataset is not directly accessed, and only the distribution is utilized.
  \item The number of samples is dynamically determined
        based on the estimated accuracy without the need for predefining it.
\end{itemize}

\subsection{Algorithm}\label{sec:alg}
{%
  \begin{figure}[t]
    \centering
    \includegraphics[width=0.32\textwidth]{rlime1}
    \includegraphics[width=0.32\textwidth]{rlime2}
    \includegraphics[width=0.32\textwidth]{rlime3}
    \caption{%
      Overview of the R-LIME algorithm.
      The progression of the algorithm is illustrated from left to right.
      The solid line represents the rectangular region $A$,
      and the dashed line represents the linear approximation model $g$
      learned within $A$.
      The initial value of $A$ is an empty rule (entire input space),
      and predicates are added to $A$, reducing coverage.
      The process continues until $\Prec(A)\ge\tau$,
      at which point the rule with the maximum coverage is output.
    }
  \end{figure}
  \begin{algorithm}[p]
    \input{alg1}
  \end{algorithm}
  \begin{algorithm}[p]
    \input{alg2}
  \end{algorithm}
  \def\myidt{\hspace{\algorithmicindent}}
  \begin{algorithm}[p]
    \input{alg3}
  \end{algorithm}
  \begin{algorithm}[p]
    \input{alg4}
  \end{algorithm}
}

Our method's algorithm is mainly based on that used in Anchor.
For non-convex optimization problems like \cref{eq:main_problem},
greedy search are often used.
However,
greedy methods tend to converge to local optima, and to address this,
R-LIME utilizes beam search,
which selects multiple candidates at each iteration.
The pseudocode is shown in \cref{alg:greedy_search}.

\subsubsection{Generation of New Candidate Rules}
To generate new candidate rules,
one additional predicate is added to each of the $B$ candidate rules
selected in the previous iteration.
The pseudocode is shown in \cref{alg:generate_cands}.
$T(x)$ is the set of attribute-value pairs (predicates) that are true for $x$,
and $T(x)\setminus A$ is the set of predicates in $T(x)$ not included in rule $A$.

\subsubsection{Selection of Candidate Rules with Maximum Accuracy}
Given the set of generated candidate rules $\cands$,
the algorithm selects the top $B$ candidates with the highest accuracy.
This is treated as an optimal arm identification problem in the multi-armed bandit framework.
Each candidate rule $A_i\in\cands$ is considered an arm,
and their accuracy $\Prec(A_i)$ is treated as the distribution of rewards.
By sampling $z\sim\mathcal{D}(\cdot|A_i)$
and obtaining the reward $\mathbbm{1}_{f(z)=g_i(z)}$ for each trial,
the algorithm updates $g_i$ using the sampled perturbation vector $z$
and the pseudo-label $f(z)$ after each trial.
To efficiently select the rule (arm) with the highest accuracy,
we employ the KL-LUCB algorithm \cite{kaufmann2013information}.
The pseudocode is shown in \cref{alg:best_cands}.
For tolerance $\epsilon\in[0,1]$, the KL-LUCB algorithm guarantees below:
\begin{equation}
  P(\underset{A\in\cands}{\min}\Prec(A)\ge
  \underset{A'\in\mathcal{A}}{\min}\Prec(A')-\epsilon)\ge1-\delta.
\end{equation}
However, 
the KL-LUCB algorithm assumes that the reward distribution for each arm
remains unchanged, 
while our method updates the classifier $g_i$ with each sampling,
which may not satisfy this assumption.
This issue is discussed further in \cref{sec:reward}.

\subsubsection{Selection of the Rule with Maximum Coverage Meeting the Constraint}
To satisfy the constraint imposed by \cref{eq:const_prec}, rule $A$ needs to fulfill the lower bound constraint:
\begin{equation}
  \Prec_{l}(A,\delta)>\tau
  \label{eq:stop_condition}
\end{equation}
where $\Prec_{l}(A,\delta)$ is the lower limit of
the $100(1-\delta)$\% confidence interval for $\Prec(A)$.
If the received set of candidate rules, $\mathcal{A}$,
includes a rule satisfying \cref{eq:stop_condition},
the one with the maximum coverage among them is selected,
and the iteration is terminated.
If $\mathcal{A}$ does not contain any rule satisfying \cref{eq:stop_condition},
it returns \textbf{null},
and proceeds to the next iteration.
The pseudocode is presented in \cref{alg:largest_valid_cand}.

% The algorithmic procedure outlined above approximates
% the solution to the optimization problem in \cref{eq:main_problem}
% under the definition of accuracy in \cref{eq:def_prec}.

\section{Experiments}
To verify the effectiveness of the proposed method,
We compare LIME and R-LIME using a real-world dataset.

@@

The experiments utilized the recidivism dataset \cite{schmidt1988predicting}.
This dataset contains information on 9,549 prisoners released from
North Carolina prisons between July 1, 1979, and June 30, 1980.
The dataset includes 19 items such as race, gender,
presence of alcohol dependence, number of prior offenses, and recidivism.
For this experiment, 
we treated the binary classification problem of predicting
the presence or absence of recidivism (Recidivism) as the target label. 
We discretized continuous features and removed missing values,
resulting in 15 features. 
This problem setting can be considered as a case
where a machine learning model is introduced to decide parole for prisoners.
Since such decisions can have a significant impact on a person's life,
it is crucial for users to interpret the outputs of black-box models appropriately.

The dataset, after removing missing values,
was split into training data (7,639 instances) and test data (955 instances).
A random forest model with 50 trees was trained using the training data.
Subsequently, LIME and R-LIME explanations were generated
for two instances extracted from the test data (\cref{fig:instance}).
In R-LIME, logistic regression was used as the linear approximation model, 
and the distribution $\mathcal{D}$ was a multivariate normal distribution
estimated from the training data.
The beam width was set to $B=10$,
the confidence coefficient to $1-\delta=0.95$,
and the tolerance of the KL-LUCB algorithm to $\epsilon=0.05$.
Accuracy thresholds $\tau$ were set to $\tau=0.70,0.80,0.90$.
